Measuring causal influence with\\ back-to-back regression

### 2019-05-23 paper/meg_exps.tex
meg results

Forward regression performs a regularized least-squares regression $A \in \mathbb{R}^{d_x \times d_y}$ from $X$ to $Y$, and estimates the causal influence $E_{i,i}$ as $\| A_{i, :} \|^2$.
%
Backward regression performs a regularized least-squares regression $B \in \mathbb{R}^{d_y \times d_x}$ from $Y$ to $X$, and estimates the causal influence $E_{i, i}$ as the Pearson correlation between the $i$-th column of $X$ and the $i$-th column of $YB$, using a validation sample.
%
For back-to-back regression, we observed that two modifications improved its performance on MEG data.
%
First, we replace the Tikhonov regularization in the first regression by truncated SVD (performed as a maximum likelihood estimation of principal components).
%
Second, we invert the data covariances using all the data, and use these common inverses across B2BR repetitions.

@@

The three types of analyses were implemented for each subject and each time
sample independently.  To estimate the reliability of these estimates in the
population, we tested, using one-sample $t$-test whether the distribution of
the estimated contribution of each factor different from baseline ($t=-100$ms),
as $\text{t-test}(\hat E_{t=-100} - \hat E_{\tau})$, with $\tau \in $ 150ms and
400ms respectively
% =======
% \paragraph{Back-to-back regression} We adapted back-to-back regression with two
% changes. First, to avoid getting a slightly positively-biased estimate of the
% factors contributing to the MEG signal, we swap the Tikhonov regularization of
% the first regression $G$ with a truncated regularization. Truncation was fitted
%  by maximum likelihood estimate, as implemented in scikit-learn principal
%  component analysis. Second, to minimize the variance increase induced by
%  bagging, we computed the inverse regularized covariance matrices outside
%  bagging:
% \begin{equation}
%   G = (Y^T Y+\lambda_Y I)^{-1} Y_{train}^T X_{train}
% \end{equation}
% \begin{equation}
% \hat X_{test} = Y_{test} \cdot G^T
% \end{equation} where $\left(train,test\right)$ are partitions of the data.
% %
% % \subsubsection{Second-level statistics}
% %
% % The three types of analyses were implemented for each subject and each time
% % sample independently.  To estimate the reliability of these estimates in the
% % population, we tested, using one-sample $t$-test whether the distribution of
% % the estimated contribution of each factor different from baseline ($t=-100$ms),
% % as $\text{t-test}(\hat E_{t=-100} - \hat E_{\tau})$, with $\tau \in $ 150ms and
% % 400ms respectively
% >>>>>>> 1775190a9737244f84b8ace529198a443029ad8d

\subsubsection{Results}

We show all the feature coefficients as a function of time in the stacked plots,
Figure \ref{fig:megresult}. Our method confirms previous findings in
neuroscience literature. In particular, \cite{kutas2011thirty} find that word
frequency correlates with a brain wave around 400ms. Similarly,
\cite{pegado2014timing} show that letter strings correlate with early brain
responses. Both these phenomena can be observed in Figures
\ref{fig:meg_twocurves} and \ref{fig:megresult}, where we show that our
method is able to separate them more clearly than the forward regression
approach.

>>>>

Forward regression performs a regularized least-squares regression $A \in \mathbb{R}^{d_x \times d_y}$ from $X$ to $Y$, and estimates the causal influence $E_{i,i}$ as $\| A_{i, :} \|^2$.
%
Backward regression performs a regularized least-squares regression $B \in \mathbb{R}^{d_y \times d_x}$ from $Y$ to $X$, and estimates the causal influence $E_{i, i}$ as the Pearson correlation between the $i$-th column of $X$ and the $i$-th column of $YB$, using a validation sample.
%
Finally, we used the back-to-back regression as described earlier, using a bagging of 10 alphas logarithmically distributed between -5 and 5.

@@

The three types of analyses were implemented for each subject and each time
sample independently.  To estimate the reliability of these estimates in the
population, we tested, using one-sample $t$-test whether the distribution of
the estimated contribution of each factor different from baseline ($t=0$ms),
as $\text{t-test}(\hat E_{t=0} - \hat E_{\tau})$, with $\tau \in $ 150ms and
400ms respectively. We use the resulting p-values as proxy to quantify the
performance of each model.
% =======
% \paragraph{Back-to-back regression} We adapted back-to-back regression with two
% changes. First, to avoid getting a slightly positively-biased estimate of the
% factors contributing to the MEG signal, we swap the Tikhonov regularization of
% the first regression $G$ with a truncated regularization. Truncation was fitted
%  by maximum likelihood estimate, as implemented in scikit-learn principal
%  component analysis. Second, to minimize the variance increase induced by
%  bagging, we computed the inverse regularized covariance matrices outside
%  bagging:
% \begin{equation}
%   G = (Y^T Y+\lambda_Y I)^{-1} Y_{train}^T X_{train}
% \end{equation}
% \begin{equation}
% \hat X_{test} = Y_{test} \cdot G^T
% \end{equation} where $\left(train,test\right)$ are partitions of the data.
% %
% % \subsubsection{Second-level statistics}
% %
% % The three types of analyses were implemented for each subject and each time
% % sample independently.  To estimate the reliability of these estimates in the
% % population, we tested, using one-sample $t$-test whether the distribution of
% % the estimated contribution of each factor different from baseline ($t=0$ms),
% % as $\text{t-test}(\hat E_{t=0} - \hat E_{\tau})$, with $\tau \in $ 150ms and
% % 400ms respectively
% >>>>>>> 1775190a9737244f84b8ace529198a443029ad8d

\subsubsection{Results}
It is a well established fact that the number of letters in a string flashed
onto the retina modulates brain activity from 150ms after stimulus onset
\cite{pegado2014timing}. Similarly, word frequency is known to modulate brain
activity around 400ms \cite{kutas2011thirty}. Consequently, we aimed to compare
how sensitive each modeling approach in detecting these two types of brain
responses in a large neuroimaging dataset.

In Figure \ref{fig:meg_twocurves} we show for each model the effects obtained in a
representative subject, as well as the average effects obtained across subjects.
The results show that all methods significantly detect that both word length and
word frequency modulate brain responses after the stimulus onset. However
forward regression detected word length effects (XXX) and word frequency (XXX)
effects less reliably than B2BR (length=XXX, frequency=XXX). B2BR appeared to
be less sensitive than the backward regression (mean XXX). However, it is
important to highlight that backward models may mistakenly confound the two
factors and thus double the signal-to-noise ratio. This possibility is
corroborated by the remarkably similar dynamics between the two curves.

Finally we show in Figure \ref{fig:megresult} that B2BR is sufficiently sensitive
to reliably detect a multiplicity of subtle features, such as individual letters
and part-of-speech. This methods thus promises to improve the detectability of
subtle effects.
